---
title: "Writing functions with with `dplyr`"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "default"
---


```{r setup, include=FALSE}
library(tidyverse)
library(skimr)
# chunk options ----
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      tidy = FALSE,
                      fig.path = "img/")
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
```


```{r , echo=FALSE}
# xaringan::inf_mr()
```


## Programming with `dplyr` (vignette)

The vignette, 'programming with `dplyr`' gives us a lot of information on how to use `dplyr` to write functions: 

We'll create a test data frame to play with for these functions:

```{r create-df}
DFX <- data.frame(a = rep(LETTERS[1:5], 2),
                  b = sample(LETTERS[1:2], 10, replace = TRUE),
                  c = sample(LETTERS[1:5], 10, replace = TRUE),
                  x = 1:10, 
                  y = rep(seq(2, 10, by = 2), 2), 
                  z = sample(1:10, 10, replace = FALSE))
DFX
DFY <- data.frame(A = rep(LETTERS[1:5], 2),
                  B = sample(LETTERS[1:2], 5, replace = TRUE),
                  C = sample(LETTERS[1:5], 10, replace = TRUE),
                  X = 1:10, 
                  Y = rep(seq(2, 10, by = 2), 2), 
                  Z = sample(1:10, 10, replace = FALSE))
DFY
```


## Data masking, env-variables, and indirection

This is an example of how masking works--the following code places two logical conditions in the 'row' position of the brackets. 

```{r bracket-subsetting}
DFX[DFX$a == "b" | DFX$x == 2, ]
```

`dplyr` makes this code more 'pipepable' and easier to read. 

```{r dplyr-subsetting}
DFX %>% filter(a == "b" | x == 2)
```


## `env-` vs. `data-` variables 

> `env-variables` are “programming” variables that ***live in an environment***. They are usually created with `<-`

we created the `df`s with `<-`, so it's in our environment:

```{r}
ls()
```


> `data-variables` are “statistical” variables that ***live in a data frame***. They usually come from data files (e.g. `.csv`, `.xls`), or are created manipulating existing variables.

```{r data-variables}
DFX$x
```
 
We now have an **env- variable** (the `DFX` `tibble`s), and four new `data-` variables (the columns in `DFX`s)

### Embracing 

If we want to refer to an **data- variable** in a function (i.e. refer to a column inside the 'data' we supply), we need to use the 'embrace' syntax: 

```{r var_summary}
var_summary <- function(data, var) {
  data %>%
    summarise(n = n(), 
              # the 'embrace' is enclosed in two 
              # double-brackets {{}}
              min = min({{ var }}), 
              max = max({{ var }})
              )
}
```


We can test this with `DFY`

```{r test-var_summary}
DFY %>% 
  group_by(B) %>% 
  var_summary(Z)
```

And we can extend `{{ var }}` to include multiple columns that come from the `env-` variable, 

```{r var_group_summary}
var_group_summary <- function(data, grpvar, sumvar) {
  data %>%
        # include a group variable
    group_by({{ grpvar }}) %>% 
    summarise(n = n(), 
              # and a summary variable
              min = min({{ sumvar }}), 
              max = max({{ sumvar }})
              )
}
```

```{r test-var_group_summary}
DFY %>% 
    var_group_summary(data = ., grpvar = B, sumvar = Z)
```


### Embrace with `.data[[]]`

If we want to make a new variable that pastes together any six possible columns

```{r create_new_column}
create_join_column <- function(df, join_cols, new_col_name){
  # select by_vars
  tmp <- select(df, all_of(join_cols))
  # rename data 
  join_col_data <- df
  # assign new col
  join_col_data$new_col <- pmap_chr(.l = tmp, .f = paste, sep = "-")
  # rename 
  names(join_col_data)[names(join_col_data) == "new_col"] <- new_col_name
  # relocate
  join_col_data <- relocate(join_col_data, all_of(new_col_name))
  # return
  return(join_col_data)
}
```


```{r dfx_2newcol}
dfx_2newcol <- create_join_column(df = DFX, 
                                  join_cols = c("a", "b"), 
                                  new_col_name = "JOINCOL")
head(dfx_2newcol)
```

```{r dfy_2newcol}
dfy_2newcol <- create_join_column(df = DFY, 
                                  join_cols = c("A", "B"), 
                                  new_col_name = "JOINCOL")
head(dfy_2newcol)
```

```{r join-dfx_2newcol-dfy_2newcol}
inner_join(x = dfx_2newcol, y = dfy_2newcol)
```

```{r}
dfx_3newcol <- create_join_column(df = DFX, join_cols = c("a", "b", "x"), new_col_name = "JOINCOL")
dfy_3newcol <- create_join_column(df = DFY, join_cols = c("A", "B", "X"), new_col_name = "JOINCOL")
inner_join(x = dfx_3newcol, y = dfy_3newcol)
```

## Extracting metadata with `fs`

Now we want to use the `fs` package to create a function that imports data into a shiny app and extracts the metadata. We will start with a basic shinydashboard layout:

```{r metadata-app}
fs::dir_tree("apps")
```

Now we'll write a function that akes a full file path and imports the metadata:

```{r}
getwd()
fs::dir_tree("/Users/mjfrigaard/Dropbox/@projects/high-priority/@r_training_materials/programming-with-dplyr/writing-dplyr-functions/data/sas")
```


```{r}
fs::file_info("data/csv/applicants.csv", follow = TRUE)
fs::file_info("data/sas/hadley.sas7bdat", follow = TRUE)
```



